Ionization of biological molecules by multicharged ions using the stoichiometric model

### 2019-10-07 arxiv/ionmol_stoi.tex
change file names

\begin{abstract}
In the present work, we investigate the ionization of molecules of
biological interest by the impact of multicharged ions in the
intermediate to high energy range. We performed full non--perturbative
distorted--wave calculations (CDW) for thirty--six collisional systems
composed by six atomic targets: H, C, N, O, F and S --which
are the constituents of most of the DNA and biological molecules--, and
six charged projectiles (antiprotons, H, He, B, C, and O). On account
of the radiation damage caused by secondary electrons, we inspect
the energy and angular distributions of the emitted electrons.
We examine seventeen molecules: DNA and RNA bases,
DNA backbone, pyrimidines, tetrahydrofuran (THF), and several
C$_n$H$_n$ compounds. We show that the simple stoichiometric model
(SSM), which approximates the molecular ionization cross sections as a
linear combination of the atomic ones, gives reasonably good results
for complex molecules. We inspect the extensively used Toburen
scaling of the total ionization cross sections of molecules with the
number of weakly bound electrons. Based on the present atomic CDW
results, we propose new active electron numbers, which
improves significantly scaling for all the targets
and ions studied here.
The new scaling describes well the available experimental
data for proton impact, including small molecules such as H$_2$, water,
methane, and ammonia.
We perform full molecular
calculations for five nucleobases and test a modified stoichiometric
formula based on the Mulliken charge of the composite atoms. The
difference introduced by the new stoichiometric formula is less than
$3\%$, which indicates the reliability of the SSM to
deal with this type of molecules. The results of the extensive
ion--target examination included in the present study allow us to
assert that the SSM and the CDW--based scaling will be useful tools in
this area.
\end{abstract}

>>>>

\begin{abstract}
In the present work, we investigate the ionization of molecules of 
biological interest by the impact of multicharged ions in the 
intermediate to high energy range. We performed full non--perturbative 
distorted--wave calculations (CDW) for thirty--six collisional systems 
composed by six atomic targets: H, C, N, O, F and S --which are the 
constituents of most of the DNA and biological molecules-- and six 
charged projectiles (antiprotons, H, He, B, C, and O). On account of 
the radiation damage caused by secondary electrons, we inspect the 
energy and angular distributions of the emitted electrons from the 
atomic targets. We examine seventeen molecules: DNA and RNA bases, 
DNA backbone, pyrimidines, tetrahydrofuran (THF), and C$_n$H$_n$ 
compounds. We show that the simple stoichiometric model (SSM), which 
approximates the molecular ionization cross sections as a linear 
combination of the atomic ones, gives reasonably good results for 
complex molecules. We also inspect the extensively used Toburen scaling 
of the total ionization cross sections of molecules with the number of 
weakly bound electrons. Based on the atomic CDW results, we 
propose new active electron numbers, which improves significantly 
scaling for all the targets and ions studied here in the intermediate 
to the high energy region. The new scaling describes well the available 
experimental data for proton impact, including small molecules. We 
perform full molecular calculations for five nucleobases and test a 
modified stoichiometric formula based on the Mulliken charge of the 
composite atoms. The difference introduced by the new stoichiometric 
formula is less than 3\%, which indicates the reliability of the SSM to 
deal with this type of molecules. The results of the extensive 
ion--target examination included in the present study allow us to assert 
that the SSM and the CDW--based scaling will be useful tools in this area.
\end{abstract}
